Multifidelity Approximate Bayesian Computation with Sequential Monte Carlo Parameter Sampling

### 2020-01-27 paper/article.tex
Flexible filenames into directory
Change location of figures to same working directory
Use bibtex output directly
Ruth's changes done

One widely-acknowledged weakness of rejection sampling ABC is the potential for wasted computational effort. 
Rejection sampling depends on repeated sampling from the prior. 
However, for $p_\epsilon(\theta~|~\obs y)$ to be a good approximation to the likelihood, $p(\theta~|~\obs y)$, we require a relatively small value of $\epsilon$.
This means that, in regions of low likelihood, many simulations will produce $y \notin \Omega_\epsilon$ and so will receive zero weight.
This often leads to poor performance: a large computational burden that produces many rejections.
In this setting, \emph{importance sampling} is one adaptation of \Cref{ABC:Rejection} commonly used to improve its performance.
It works by carrying out fewer simulations in regions of low likelihood~\cite{Owen2013,Sisson2018} by proposing parameters according to an importance distribution defined on $\Theta$.

@@

\begin{lemma}
\label{lemma:MC}
Consider the Monte Carlo sample $\{ \theta_n, w_n \}_{n=1}^N$ generated by an ABC algorithm where each $\theta_n \sim \hat q$ is independently generated from the importance distribution, $\hat q$, and each weight, $w_n$, is randomly generated conditionally on $\theta_n$.
Suppose that there exists a positive constant $\lambda$ such that
\begin{equation}
\label{eq:ABCCondition}
\mathbb E(w~|~\theta) \hat q(\theta) = \lambda f_\epsilon(\obs y~|~\theta) \pi(\theta),
\end{equation}
where $f_\epsilon$ is the ABC approximation to the likelihood and where $\pi(\cdot)$ is the prior.
Then, for arbitrary functions $F$ on $\Theta$, the mean squared error (MSE),
\[
\mathbb E\left( \left( \bar F - \mathbb E_{p_\epsilon} \left(F(\theta) \right) \right)^2 \right) \rightarrow 0,
\]
between $\mathbb E_{p_\epsilon}(F(\theta))$ and $\bar F$, defined in \Cref{eq:estimate}, decays as $N \rightarrow \infty$ on the order of $1/N$.
Thus, the Monte Carlo sample is from the ABC posterior $p_\epsilon(\theta~|~\obs y)$.
\end{lemma}
\begin{proof}
For an arbitrary pair $(w, \theta)$ from the Monte Carlo sample, the expected value of $w F(\theta)$ can be conditioned on $\theta$ to give 
\begin{align*}
\mathbb E(w F(\theta)) 
&= \int F(\theta) \mathbb E(w~|~\theta) \hat q(\theta) ~\mathrm d\theta 
\\ 
&= \lambda \int F(\theta) f_\epsilon(\obs y~|~\theta) \pi(\theta) ~\mathrm d\theta 
\\
&= \lambda Z \mathbb E_{p_\epsilon}(F(\theta)).
\end{align*}
For the particular case where $F \equiv 1$, it follows that $\mathbb E(w) = \lambda Z$.

Using the delta method~\cite{Seltman2018}, we can therefore write both the bias and variance of the ratio in \Cref{eq:estimate}, to leading order, as
\begin{subequations}
\label{eq:estimate:MSE}
\begin{align}
\mathbb E \left( \bar F \right) - \mathbb E_{p_\epsilon}(F(\theta))
&= 
\frac{1}{N} \frac{\mathbb E \left( w^2 \left[ \mathbb E_{p_\epsilon}(F(\theta)) - F \left(\theta\right) \right] \right)}{\mathbb E(w)^2} + O(1/N^2),
\label{eq:estimate:bias}
\\
\mathrm{Var} \left( \bar F \right)
&=
\frac{1}{N} \frac{\mathbb E \left( w^2 \left[ \mathbb E_{p_\epsilon}(F(\theta)) - F\left(\theta\right) \right]^2 \right) }{\mathbb E(w)^2} + O(1/N^2).
\label{eq:estimate:variance}
\end{align}
\end{subequations}
Using the standard decomposition of the MSE into the sum of the variance and squared bias, it follows that, to leading order, the MSE is dominated by the variance term in \Cref{eq:estimate:variance} as $N \rightarrow \infty$.
Therefore the estimate in \Cref{eq:estimate} is consistent in the limit of large sample size, with convergence in MSE on the order of $1/N$. 
\end{proof}

@@

One important aspect of this algorithm is the method for sampling from each importance distribution $\hat q_t$, with density proportional to $q_t$, as defined in \Cref{eq:importance}.
The following procedure generates $\theta^\star \sim \hat q_{t+1}(\cdot)$:
\begin{enumerate}
\item Choose a parameter $\theta^{\star\star}$ from the set $\{ \theta_n^{(t)} \}_{n=1}^{N_t}$, with probabilities proportional to the weights $\{ w_n^{(t)} \}_{n=1}^{N_t}$.
\item Perturb the parameter according to a perturbation kernel $K_t$, producing $\theta^{\star} \sim K_{t}(\cdot~|~\theta^{\star\star})$.
\item If $\pi(\theta)>0$, then return $\theta^{\star}$; else return to 1.
\end{enumerate}
Note that the ABC-SMC algorithm requires the specification of a threshold sequence $\epsilon_t$, the stopping conditions $S_t$, and the perturbation kernels $K_t(\cdot~|~\theta)$ for $t=1,\dots,T$.
We will not re-examine these aspects of ABC-SMC in detail in this paper, and will implement ABC-SMC using established techniques to select $\epsilon_t$, $S_t$ and $K_t$~\cite{Toni2009,DelMoral2012,Beaumont2009,Filippi2013}.

\subsection{Multifidelity ABC}
\label{s:MFABC}
Importance sampling and SMC approaches can improve the performance of ABC by exploring parameter space more efficiently, so that simulation time is not wasted in regions of low likelihood.
An orthogonal approach to improving ABC is to attempt to avoid computationally expensive simulations, where possible, by relying on the \emph{multifidelity} framework~\cite{Prescott2020}.
We term the model of interest, $f(\cdot~|~\theta)$, which maps parameter space $\Theta$ to an output space $\mathcal Y$, as the high-fidelity model, .
We now assume that, in addition, there is a low-fidelity (i.e. approximate) model, $\tilde f(\cdot~|~\theta)$, of the same physical system with the same parameter space, $\Theta$.
The simulations of this model are denoted $\tilde y \sim \tilde f(\cdot~|~\theta)$, taking values in the output space $\tilde{\mathcal Y}$, which may differ from $\mathcal Y$.
Importantly, we assume that the low-fidelity model is computationally cheaper, in the sense that simulations $\tilde y \sim \tilde f(\cdot~|~\theta)$ of the low-fidelity model incur less computational burden than simulations $y \sim f(\cdot~|~\theta)$ of the high-fidelity model.

@@

In general, the models $f(\cdot~|~\theta)$ and $\tilde f(\cdot~|~\theta)$ can be simulated independently for a given $\theta$.
Then, if the low-fidelity model is a good approximation to the high-fidelity model, the outputs $y$ and $\tilde y$ will be near, in some sense.
However, we will also allow for \emph{coupling} between the two models, 
by which we mean that a simulation from the high-fidelity model can be conditioned on a low-fidelity simulation, $\tilde y \sim \tilde f(\cdot~|~\theta)$~\cite{Giles2008}.
The coupling approach requires us to specify the \emph{coupled} high-fidelity model, with probability distribution $f(\cdot~|~\tilde y, \theta)$.
The key condition for a valid coupled high-fidelity model is that
\begin{equation}
\label{eq:coupling}
 \int_{\tilde{\mathcal{Y}}} f(y ~|~ \tilde y, \theta) \tilde f(\tilde y~|~\theta) ~\mathrm d\tilde y = f(y ~|~ \theta).
\end{equation}
The benefit of this approach is that, with a judicious choice of coupling method, a simulation $y \sim f(\cdot~|~\tilde y, \theta)$ from the coupled high-fidelity model is more closely correlated to $\tilde y$ than an independent simulation $y \sim f(\cdot~|~\theta)$ from the original high-fidelity model.
Furthermore, simulations of $y \sim f(\cdot~|~\tilde y, \theta)$ may be less computationally burdensome than simulating $y \sim f(\cdot~|~\theta)$.

As an example, suppose the high-fidelity model is a Markovian stochastic dynamical system on the time horizon $t \in [0,T]$, and that the low-fidelity model is the same system on $t \in [0,\tau]$ for $\tau<T$~\cite{Prangle2016}.
The low-fidelity and high-fidelity models can clearly be simulated independently on $[0,\tau]$ and $[0,T]$, respectively, and the low-fidelity simulation will, on average, be less computationally expensive than the high-fidelity simulation.
However, suppose we have generated a simulation, $\tilde y(t)$, of the low-fidelity model, which is a trajectory over $t \in [0,\tau]$.
In this case, a natural coupling is to deterministically assign $y(t)$ to have the same trajectory as $\tilde y(t)$ over $t \in [0, \tau]$, and then complete the simulation for $y(t)$ over $t \in [\tau, T]$ from the initial condition $\tilde y(\tau)$.
The coupled high-fidelity simulation, $y$, generated conditionally on $\tilde y$, incurs further simulation cost beyond that of $\tilde y$, but this is less than that incurred by generating an independent (i.e. uncoupled) high-fidelity simulation.
Many other possible couplings exist for different multifidelity models, often involving shared random noise processes, and methods for coupling are currently an area of active research~\cite{Croci2018,Lester2018,Prescott2020}.

@@

\begin{proposition}
\label{MFABCValidity}
The weighted Monte Carlo sample $\{ \theta_n, w_n\}$ returned by \Cref{MFABC:Rejection} is from the ABC posterior $p_\epsilon(\theta~|~\obs y)$.
\end{proposition}
\begin{proof}
By \Cref{lemma:MC}, we need to show that \Cref{eq:ABCCondition} holds.
Here, $\hat q = \pi$, and performing the triple integral in the order $u$, $\tilde y$, then $y$ gives the equality
\begin{align*}
\mathbb E(w(\tilde y, u, y)~|~\theta) 
&= \iiint w(\tilde y, u, y) f(y~|~\tilde y, \theta) \tilde f(\tilde y~|~\theta) ~\mathrm du ~\mathrm d\tilde y ~\mathrm dy 
\\
&= \iint \mathbb I(y \in \Omega_\epsilon) f(y~|~\tilde y, \theta) \tilde f(\tilde y~|~\theta) ~\mathrm d\tilde y ~\mathrm dy
\\
&= \int \mathbb I(y \in \Omega_\epsilon) f(y~|~\theta) ~\mathrm dy
\\
&= f_\epsilon(\obs y~|~\theta).
\end{align*}
Hence \Cref{eq:ABCCondition} holds for $w_n = w(\tilde y_n, u_n, y_n)$ given by \Cref{eq:w_mf}, and the result follows.
\end{proof}

Note that smaller values of continuation probability, $\alpha$, lead to greater savings in computational burden, since high-fidelity simulations, $y_n$, are generated less often.
However, it is possible to show that this comes at the cost of an increase in the MSE of the estimator as given in \Cref{eq:estimate}, $\bar F \approx \mathbb E_{p_\epsilon}(F(\cdot))$, of an arbitrary function $F(\cdot)$ on $\Theta$.
We now describe the trade-off between reducing computational burden and increasing MSE in more detail.

@@

We can express computational burden more explicitly in terms of simulation costs.
For concreteness, we define $T_{\mathrm{total}} = \sum_n T_n$ to be the sum of the elapsed times, $T_n = T(w_n)$, taken to calculate each $w_n$ given $\theta_n$.
Then $\mathbb E(T_{\mathrm{total}}) = N \mathbb E(T)$, where the random variable $T$ denotes the time taken to generate a single weight $w$.
Therefore,
\[
\mathbb E(T_{\mathrm{total}}) 
\approx
\frac{\mathrm{Var}_{p_\epsilon} (F(\theta)) / \mathrm{MSE} }{Z / \mathbb E(T)},
\]
gives an estimate of the expected simulation time required to construct a Monte Carlo estimate, $\bar F \approx \mathbb E_{p_\epsilon}(F(\theta))$ using \Cref{ABC:Rejection} (ABC-RS) with a target MSE value.
In this formulation of computational cost, we will interpret the ratio $Z/\mathbb E(T)$ as the theoretical efficiency of \Cref{ABC:Rejection}, according to \Cref{def:efficiency:th} below.

@@

Since the weights, $w_n$, are now calculated using the multifidelity weight in \Cref{eq:qw_mf}, there is a positive probability that there exist $w_n < 0$.
These are generated whenever $\tilde y_n \in \Omega_\epsilon$, but where we also simulate the high-fidelity model and produce $y_n \notin \Omega_\epsilon$.
This leads to two problems with the existing definition of the importance distribution in \Cref{eq:importance}.
First, there is no longer a guarantee that $q_{t+1}(\theta) > 0$ whenever $\pi(\theta)>0$, even when the perturbation kernels, $K_t$, have infinite support.
Secondly, even if the $K_t$ could be chosen to guarantee $q_{t+1}(\theta)>0$, the method for sampling from the importance distribution, $\hat q_{t+1} \propto q_{t+1}$, is no longer valid: 
we cannot choose $\theta^{\star \star}$ from the set of $\{ \theta_n^{(t)} \}_{n=1}^{N_t}$ with probabilities proportional to $\{ w_n^{(t)} \}_{n=1}^{N_t}$, because any $w_n^{(t)} < 0$ cannot be a probability.

To overcome the first issue, we define a new importance distribution based on a parameter, $0<\delta<1$, that uses the technique of defensive importance sampling~\cite{Owen2013} and enforces positivity.
Dropping the index $t$ for notational simplicity, we define the two functions
\begin{subequations}
\label{eq:DefensiveImportance}
\begin{align}
 r(\theta) &= \delta \pi(\theta) + (1-\delta) \max\left( 0, q(\theta) \right),
 \\
 q(\theta) &= \begin{cases}
                      \sum_{n=1}^{N} w_n K(\theta ~|~ \theta_n) \big/ \sum_{m=1}^{N} w_m & \pi(\theta)>0, 
                      \\
                      0 & \text{else,}
                    \end{cases}
\end{align}
\end{subequations}
where $q$ has exactly the same form given in \Cref{eq:importance}.
Clearly, $r(\theta) \geq \delta \pi(\theta)$ is positive for all $\theta$ such that $\pi(\theta)>0$, and it therefore induces a valid importance distribution $\hat r(\theta) = r(\theta)/Z_r$.
Furthermore, the parameter $\delta$ is useful because, by using $\hat r$ as the importance distribution in \Cref{eq:qw_mf}, we can enforce an upper bound of $\pi(\theta)/r(\theta) \leq 1/\delta$ on the ratio scaling the multifidelity weight.
The trade-off is that using $\delta>0$ will redistribute some of the importance density away from regions of high likelihood (where $q$ is large and positive) back onto the prior distribution.

@@

To illustrate this sampling procedure, we let $f$ and $g$, respectively, be the probability density functions of $\mathrm{Beta}(1,2)$ and $\mathrm{Beta}(2,2)$ random variables.
Choosing $\mu = 3$, \Cref{fig:negative_mixture} shows that $h(x)<0$ on $x>0.5$.
We applied the sampling procedure in \Cref{negmix} for each of $\phi_0(x) \equiv 0$ and $\phi_1(x) = 0.3f(x)$ to produce $20,000$ sample values from $\hat h_{\phi_0}$ and $\hat h_{\phi_1}$, and plotted their histograms.
The acceptance probabilities in step (3) of \Cref{negmix} were $\max(0, 1-\alpha)$ and $\max(0.1, 1 - \alpha)$, respectively.
The results in \Cref{fig:negative_mixture} demonstrate that we can sample from a negative mixture and simultaneously specify a positive lower bound on the resulting density.

\begin{figure}
\centering
\begin{subfigure}[b]{0.48\textwidth}
 \includegraphics[width=\textwidth]{negative_mixture_A.png}
 \caption{Sampling from $\hat h_{\phi_0} \propto h_{\phi_0} = \max(0,h)$.}
 \label{fig:negative_mixture_A}
\end{subfigure}
~
\begin{subfigure}[b]{0.48\textwidth}
 \includegraphics[width=\textwidth]{negative_mixture_B.png}
 \caption{Sampling from $\hat h_{\phi_1} \propto h_{\phi_1} = \max(0.3f, h)$.}
 \label{fig:negative_mixture_B}
\end{subfigure}
\caption{Sampling from a negative mixture. Here, $f$ and $g$ are the probability density functions of $\mathrm{Beta}(1,2)$ and $\mathrm{Beta}(2,2)$ distributions.}
\label{fig:negative_mixture}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=0.95\columnwidth]{negative_mixture_q.png}
\caption{
Sampling from \Cref{NegativeResample} for $\hat r$ given by \Cref{eq:DefensiveImportance}, formed from four parameter values $\theta_n \in \{-0.5, 0, 0.5, 1\}$ with corresponding weights $w_n \in \{2, 1, 1, -0.75\}$. 
The perturbation kernels $K(\cdot~|~\theta_n)$ are Gaussian, centred at $\theta_n$ with standard deviation $0.4$.
The weighted perturbation kernels are displayed using dotted blue lines. 
The prior is $\pi \sim \mathrm{Unif}(-2, 2)$, and the defensive importance sampling parameter is $\delta = 0.1$.
}
\label{fig:negative_mixture_q}
\end{figure}

@@

Recall that we are aiming to use this method in the context of sequential Monte Carlo.
In the sequential importance sampling approach discussed in \Cref{s:SMC,s:MF-ABC-SIS}, we use a Monte Carlo sample from generation $t$ to construct the importance distribution, $\hat q_{t+1}$ or $\hat r_{t+1}$, for use in generation $t+1$.
The proposed approach to constructing continuation probabilities is the same: we will use the Monte Carlo sample built in generation $t$ to also produce continuation probabilities, $\eta_1$ and $\eta_2$, with which we define the continuation probability $\alpha_{t+1}$ to use at generation $t+1$.
The following definition specifies how to calculate approximations of the quantities in \Cref{eq:PhiComponents} using an existing Monte Carlo sample.
These approximations can then be substituted into \Cref{eq:etastar:unbounded,eq:etastar:boundary}.
Hence, we can estimate the optimal continuation probabilities as given by \Cref{etastar}.

>>>>

One widely-acknowledged weakness of rejection sampling ABC is the potential for wasted computational effort. 
Rejection sampling depends on repeated sampling from the prior. 
However, for $p_\epsilon(\theta~|~\obs y)$ to be a good approximation to the likelihood, $p(\theta~|~\obs y)$, we require a relatively small value of $\epsilon$.
This means that, in regions of low likelihood, many simulations will produce $y \notin \Omega_\epsilon$ and so will receive zero weight.
This often leads to poor performance: a large computational burden that produces many rejections.
In this setting, \emph{importance sampling} is one adaptation of \Cref{ABC:Rejection} commonly used to improve its performance.
It works by carrying out fewer simulations in regions of low likelihood~\cite{Sisson2018,Owen2013} by proposing parameters according to an importance distribution defined on $\Theta$.

@@

\begin{lemma}
\label{lemma:MC}
Consider the Monte Carlo sample $\{ \theta_n, w_n \}_{n=1}^N$ generated by an ABC algorithm where each $\theta_n \sim \hat q$ is independently generated from the importance distribution, $\hat q$, and each weight, $w_n$, is randomly generated conditionally on $\theta_n$.
Suppose that there exists a positive constant $\lambda$ such that
\begin{equation}
\label{eq:ABCCondition}
\mathbb E(w~|~\theta) \hat q(\theta) = \lambda f_\epsilon(\obs y~|~\theta) \pi(\theta),
\end{equation}
where $f_\epsilon$ is the ABC approximation to the likelihood and where $\pi(\cdot)$ is the prior.
Then, for arbitrary functions $F$ on $\Theta$, the mean squared error (MSE),
\[
\mathbb E\left( \left( \bar F - \mathbb E_{p_\epsilon} \left(F(\theta) \right) \right)^2 \right),
\]
between $\mathbb E_{p_\epsilon}(F(\theta))$ and $\bar F$, defined in \Cref{eq:estimate}, decays as $N \rightarrow \infty$ on the order of $1/N$.
Thus, the Monte Carlo sample is from the ABC posterior $p_\epsilon(\theta~|~\obs y)$.
\end{lemma}
\begin{proof}
For an arbitrary pair $(w, \theta)$ from the Monte Carlo sample, the expected value of $w F(\theta)$ can be conditioned on $\theta$ to give 
\begin{align*}
\mathbb E(w F(\theta)) 
&= \int F(\theta) \mathbb E(w~|~\theta) \hat q(\theta) ~\mathrm d\theta 
\\ 
&= \lambda \int F(\theta) f_\epsilon(\obs y~|~\theta) \pi(\theta) ~\mathrm d\theta 
\\
&= \lambda Z \mathbb E_{p_\epsilon}(F(\theta)).
\end{align*}
For the particular case where $F \equiv 1$, it follows that $\mathbb E(w) = \lambda Z$.

Using the delta method~\cite{Seltman2018}, we can therefore write both the bias and variance of the ratio in \Cref{eq:estimate}, to leading order, as
\begin{subequations}
\label{eq:estimate:MSE}
\begin{align}
\mathrm{Bias}(\bar F) = \mathbb E \left( \bar F \right) - \mathbb E_{p_\epsilon}(F(\theta))
&= 
\frac{1}{N} \frac{\mathbb E \left( w^2 \left[ \mathbb E_{p_\epsilon}(F(\theta)) - F \left(\theta\right) \right] \right)}{\mathbb E(w)^2} + O(1/N^2),
\label{eq:estimate:bias}
\\
\mathrm{Var} \left( \bar F \right)
&=
\frac{1}{N} \frac{\mathbb E \left( w^2 \left[ \mathbb E_{p_\epsilon}(F(\theta)) - F\left(\theta\right) \right]^2 \right) }{\mathbb E(w)^2} + O(1/N^2).
\label{eq:estimate:variance}
\end{align}
\end{subequations}
Using the standard decomposition of the MSE into the sum of the variance and squared bias, it follows that, to leading order, the MSE is dominated by the variance term given in \Cref{eq:estimate:variance} as $N \rightarrow \infty$.
Therefore the estimate in \Cref{eq:estimate} is consistent in the limit of large sample size, with convergence in MSE on the order of $1/N$. 
\end{proof}

@@

One important aspect of this algorithm is the method for sampling from each importance distribution $\hat q_t$, with density proportional to $q_t$, as defined in \Cref{eq:importance}.
The following procedure generates $\theta^\star \sim \hat q_{t+1}(\cdot)$:
\begin{enumerate}
\item Choose a parameter $\theta^{\star\star}$ from the set $\{ \theta_n^{(t)} \}_{n=1}^{N_t}$, with probabilities proportional to the weights $\{ w_n^{(t)} \}_{n=1}^{N_t}$.
\item Perturb the parameter according to a perturbation kernel $K_t$, producing $\theta^{\star} \sim K_{t}(\cdot~|~\theta^{\star\star})$.
\item If $\pi(\theta)>0$, then return $\theta^{\star}$; else return to 1.
\end{enumerate}
Note that the ABC-SMC algorithm requires the specification of a threshold sequence $\epsilon_t$, the stopping conditions $S_t$, and the perturbation kernels $K_t(\cdot~|~\theta)$ for $t=1,\dots,T$~\cite{Sisson2018}.
We will not re-examine these aspects of ABC-SMC in detail in this paper, and will implement ABC-SMC using established techniques to select $\epsilon_t$, $S_t$ and $K_t$~\cite{Toni2009,DelMoral2012,Beaumont2009,Filippi2013}.

\subsection{Multifidelity ABC}
\label{s:MFABC}
Importance sampling and SMC approaches can improve the performance of ABC by exploring parameter space more efficiently, so that simulation time is not wasted in regions of low likelihood.
An orthogonal approach to improving ABC is to attempt to avoid computationally expensive simulations, where possible, by relying on the \emph{multifidelity} framework~\cite{Prescott2020}.
We term the model of interest, $f(\cdot~|~\theta)$, which maps parameter space $\Theta$ to an output space $\mathcal Y$, as the high-fidelity model.
We now assume that, in addition, there is a low-fidelity (i.e. approximate) model, $\tilde f(\cdot~|~\theta)$, of the same physical system with the same parameter space $\Theta$.
The simulations of this model are denoted $\tilde y \sim \tilde f(\cdot~|~\theta)$, taking values in the output space $\tilde{\mathcal Y}$, which may differ from $\mathcal Y$.
Importantly, we assume that the low-fidelity model is computationally cheaper, in the sense that simulations $\tilde y \sim \tilde f(\cdot~|~\theta)$ of the low-fidelity model incur less computational burden than simulations $y \sim f(\cdot~|~\theta)$ of the high-fidelity model.

@@

In general, the models $f(\cdot~|~\theta)$ and $\tilde f(\cdot~|~\theta)$ can be simulated independently for a given $\theta$.
Then, if the low-fidelity model is a good approximation to the high-fidelity model, the outputs $y$ and $\tilde y$ will be near, in some sense.
However, we will also allow for \emph{coupling} between the two models, 
by which we mean that a simulation from the high-fidelity model can be conditioned on a low-fidelity simulation, $\tilde y \sim \tilde f(\cdot~|~\theta)$~\cite{Giles2008}.
The coupling approach requires us to specify the \emph{coupled} high-fidelity model, with probability distribution $f(\cdot~|~\tilde y, \theta)$.
The key condition for a valid coupled high-fidelity model is that
\begin{equation}
\label{eq:coupling}
 \int_{\tilde{\mathcal{Y}}} f(y ~|~ \tilde y, \theta) \tilde f(\tilde y~|~\theta) ~\mathrm d\tilde y = f(y ~|~ \theta).
\end{equation}
The benefit of this approach is that, with a judicious choice of coupling method, a simulation $y \sim f(\cdot~|~\tilde y, \theta)$ from the coupled high-fidelity model is more closely correlated to $\tilde y$ than an independent simulation $y \sim f(\cdot~|~\theta)$ from the original high-fidelity model.
Furthermore, simulations of $y \sim f(\cdot~|~\tilde y, \theta)$ may be less computationally burdensome than simulations of $y \sim f(\cdot~|~\theta)$.

As an example, suppose the high-fidelity model is a Markovian stochastic dynamical system on the time horizon $t \in [0,T]$, and that the low-fidelity model is the same system on $t \in [0,\tau]$ for $\tau<T$~\cite{Prangle2016}.
The low-fidelity and high-fidelity models can clearly be simulated independently on $[0,\tau]$ and $[0,T]$, respectively, and the low-fidelity simulation will, on average, be less computationally expensive than the high-fidelity simulation.
However, suppose we have generated a simulation, $\tilde y(t)$, of the low-fidelity model, which is a trajectory over $t \in [0,\tau]$.
In this case, a natural coupling is to deterministically assign $y(t)$ to have the same trajectory as $\tilde y(t)$ over $t \in [0, \tau]$, and then complete the simulation for $y(t)$ over $t \in [\tau, T]$ from the initial condition $\tilde y(\tau)$.
The coupled high-fidelity simulation, $y$, generated conditionally on $\tilde y$, incurs further simulation cost beyond that of $\tilde y$, but this is less than that incurred by generating an independent (i.e. uncoupled) high-fidelity simulation.
Many other possible couplings exist for different multifidelity models, often involving shared random noise processes, and methods for coupling are currently an area of active research~\cite{Prescott2020,Croci2018,Lester2018}.

@@

\begin{proposition}
\label{MFABCValidity}
The weighted Monte Carlo sample $\{ \theta_n, w_n\}$ returned by \Cref{MFABC:Rejection} is from the ABC posterior $p_\epsilon(\theta~|~\obs y)$.
\end{proposition}
\begin{proof}
By \Cref{lemma:MC}, we need to show that \Cref{eq:ABCCondition} holds.
Here $\hat q = \pi$ and performing the triple integral in the order $u$, $\tilde y$, then $y$ gives the equality
\begin{align*}
\mathbb E(w(\tilde y, u, y)~|~\theta) 
&= \iiint w(\tilde y, u, y) f(y~|~\tilde y, \theta) \tilde f(\tilde y~|~\theta) ~\mathrm du ~\mathrm d\tilde y ~\mathrm dy 
\\
&= \iint \mathbb I(y \in \Omega_\epsilon) f(y~|~\tilde y, \theta) \tilde f(\tilde y~|~\theta) ~\mathrm d\tilde y ~\mathrm dy
\\
&= \int \mathbb I(y \in \Omega_\epsilon) f(y~|~\theta) ~\mathrm dy
\\
&= f_\epsilon(\obs y~|~\theta).
\end{align*}
Hence \Cref{eq:ABCCondition} holds for $w_n = w(\tilde y_n, u_n, y_n)$ given by \Cref{eq:w_mf}, and the result follows.
\end{proof}

Note that smaller values of the continuation probability, $\alpha$, lead to greater savings in computational burden, since high-fidelity simulations, $y_n$, are generated less often.
However, it is possible to show that this comes at the cost of an increase in the MSE of the estimator, $\bar F \approx \mathbb E_{p_\epsilon}(F(\cdot))$, (as given in \Cref{eq:estimate}) of an arbitrary function $F(\cdot)$ on $\Theta$.
We now describe in more detail the trade-off between reducing the computational burden and increasing the MSE.

@@

We can express the computational burden more explicitly in terms of simulation costs.
For concreteness, we define $T_{\mathrm{total}} = \sum_n T_n$ to be the sum of the elapsed times, $T_n = T(w_n)$, taken to calculate each $w_n$ given $\theta_n$.
Then $\mathbb E(T_{\mathrm{total}}) = N \mathbb E(T)$, where the random variable $T$ denotes the time taken to generate a single weight $w$.
Therefore,
\[
\mathbb E(T_{\mathrm{total}}) 
\approx
\frac{\mathrm{Var}_{p_\epsilon} (F(\theta)) / \mathrm{MSE} }{Z / \mathbb E(T)},
\]
gives an estimate of the expected simulation time required to construct a Monte Carlo estimate, $\bar F \approx \mathbb E_{p_\epsilon}(F(\theta))$, using \Cref{ABC:Rejection} (ABC-RS) with a target MSE value.
In this formulation of computational cost, we will interpret the ratio $Z/\mathbb E(T)$ as the theoretical efficiency of \Cref{ABC:Rejection}, according to \Cref{def:efficiency:th} below.

@@

Since the weights, $w_n$, are now calculated using the multifidelity weight in \Cref{eq:qw_mf}, there is a positive probability that there exist $w_n < 0$.
Negative values of $w_n$ are generated whenever $\tilde y_n \in \Omega_\epsilon$, but where we also simulate the high-fidelity model and produce $y_n \notin \Omega_\epsilon$.
This leads to two problems with the existing definition of the importance distribution in \Cref{eq:importance}.
First, there is no longer a guarantee that $q_{t+1}(\theta) > 0$ whenever $\pi(\theta)>0$, even when the perturbation kernels, $K_t$, have infinite support.
Secondly, even if the $K_t$ could be chosen to guarantee $q_{t+1}(\theta)>0$, the method for sampling from the importance distribution, $\hat q_{t+1} \propto q_{t+1}$, is no longer valid: 
we cannot choose $\theta^{\star \star}$ from the set of $\{ \theta_n^{(t)} \}_{n=1}^{N_t}$ with probabilities proportional to $\{ w_n^{(t)} \}_{n=1}^{N_t}$, because any $w_n^{(t)} < 0$ cannot define a probability.

To overcome the first issue, we define a new importance distribution based on a parameter $0<\delta<1$ that uses the technique of defensive importance sampling~\cite{Owen2013} and enforces positivity.
Dropping the index $t$ for notational simplicity, we define the two functions
\begin{subequations}
\label{eq:DefensiveImportance}
\begin{align}
 r(\theta) &= \delta \pi(\theta) + (1-\delta) \max\left( 0, q(\theta) \right),
 \\
 q(\theta) &= \begin{cases}
                      \sum_{n=1}^{N} w_n K(\theta ~|~ \theta_n) \big/ \sum_{m=1}^{N} w_m & \pi(\theta)>0, 
                      \\
                      0 & \text{else,}
                    \end{cases}
\end{align}
\end{subequations}
where $q$ has exactly the same form given in \Cref{eq:importance}.
Clearly, $r(\theta) \geq \delta \pi(\theta)$ is positive for all $\theta$ such that $\pi(\theta)>0$, and it therefore induces a valid importance distribution $\hat r(\theta) = r(\theta)/Z_r$.
Furthermore, the parameter $\delta$ is useful because, by using $\hat r$ as the importance distribution in \Cref{eq:qw_mf}, we can enforce an upper bound of $\pi(\theta)/r(\theta) \leq 1/\delta$ on the ratio scaling the multifidelity weight.
The trade-off is that using $\delta>0$ will redistribute some of the importance density away from regions of high likelihood (where $q$ is large and positive) back onto the prior distribution.

@@

To illustrate this sampling procedure, we let $f$ and $g$, respectively, be the probability density functions of $\mathrm{Beta}(1,2)$ and $\mathrm{Beta}(2,2)$ random variables.
Choosing $\mu = 3$, \Cref{fig:negative_mixture} shows that $h(x)<0$ on $x>0.5$.
We applied the sampling procedure in \Cref{negmix} for each of $\phi_0(x) \equiv 0$ and $\phi_1(x) = 0.3f(x)$ to produce $20,000$ sample values from $\hat h_{\phi_0}$ and $\hat h_{\phi_1}$, and plotted their histograms.
The acceptance probabilities in step (3) of \Cref{negmix} were $\max(0, 1-\beta)$ and $\max(0.1, 1 - \beta)$, respectively.
The results in \Cref{fig:negative_mixture} demonstrate that we can sample from a negative mixture and simultaneously specify a positive lower bound on the resulting density.

\begin{figure}
\centering
\begin{subfigure}[b]{0.48\textwidth}
 \includegraphics[width=\textwidth]{negative_mixture_A}
 \caption{Sampling from $\hat h_{\phi_0} \propto h_{\phi_0} = \max(0,h)$.}
 \label{fig:negative_mixture_A}
\end{subfigure}
~
\begin{subfigure}[b]{0.48\textwidth}
 \includegraphics[width=\textwidth]{negative_mixture_B}
 \caption{Sampling from $\hat h_{\phi_1} \propto h_{\phi_1} = \max(0.3f, h)$.}
 \label{fig:negative_mixture_B}
\end{subfigure}
\caption{Sampling from a negative mixture. Here, $f$ and $g$ are the probability density functions of $\mathrm{Beta}(1,2)$ and $\mathrm{Beta}(2,2)$ distributions.}
\label{fig:negative_mixture}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=0.95\columnwidth]{negative_mixture_q}
\caption{
Sampling from \Cref{NegativeResample} for $\hat r$ given by \Cref{eq:DefensiveImportance}, formed from four parameter values $\theta_n \in \{-0.5, 0, 0.5, 1\}$ with corresponding weights $w_n \in \{2, 1, 1, -0.75\}$. 
The perturbation kernels $K(\cdot~|~\theta_n)$ are Gaussian, centred at $\theta_n$ with standard deviation $0.4$.
The weighted perturbation kernels are displayed using dotted blue lines. 
The prior is $\pi \sim \mathrm{Unif}(-2, 2)$, and the defensive importance sampling parameter is $\delta = 0.1$.
}
\label{fig:negative_mixture_q}
\end{figure}

@@

Recall that we are aiming to use this method in the context of sequential Monte Carlo.
In the sequential importance sampling approach discussed in \Cref{s:SMC,s:MF-ABC-SIS}, we use a Monte Carlo sample from generation $t$ to construct the importance distribution, $\hat q_{t+1}$ or $\hat r_{t+1}$, for use in generation $t+1$.
The proposed approach to constructing continuation probabilities is the same: we will use the Monte Carlo sample built in generation $t$ to also produce the values of $\eta_1$ and $\eta_2$ defining the continuation probability $\alpha_{t+1}$ to use at generation $t+1$.
The following definition specifies how to calculate approximations of the quantities in \Cref{eq:PhiComponents} using an existing Monte Carlo sample.
These approximations can then be substituted into \Cref{eq:etastar:unbounded,eq:etastar:boundary}.
Hence, we can estimate the optimal continuation probabilities as given by \Cref{etastar}.
